\section{Conclusion}

\subsection{Future work}

\rad's hyperstatic environment solves one problem but brings with it another:
arbitrarily modifying the behaviour of previously defined functions is now
impossible. If all the behaviours that are likely to be modified in the future of
a chain are stored as functions behind refs, then those behaviours can be
updated during the lifetime of the chain using \texttt{write-ref}, and accessed
in subsequent definitions using \texttt{read-ref}. However it is likely that the
full range of desired modifications is not known at the time the chain is
created. In that case, if a core function needs to be updated, then
redefinitions of all the functions that depend on it also need to be submitted.
This wholesale redefinition of most of the environment is computationally
expensive and wasteful. In the future we would like to explore ways to mitigate
this issue; possibly allowing environment modifications as an effect.

Related to this is the lack of a module system: a system to package up common
functionality. At the moment if a user wants to use some of the functions
defined in a file \texttt{useful-fns.rad}, the easiest thing to do is to just
send all the code contained in this file to the chain, and then to submit the
definition which uses these functions. This is inefficient for several reasons:
\begin{itemize}
  \item Some of the definitions might already be in effect on the chain,
  \item Most likely not all the definitions in the file are actually needed for
    the functionality the user wants to add to a chain.
\end{itemize}
In the future we are likely to add tooling for defining a function locally, and
then extracting the minimal set of dependencies needed before submitting this to
a chain.

\rad{} is designed to be used in scenarios where the traditional way of fixing a bug---simply rewriting the program---is not available, since once inputs are accepted they cannot be taken back. Moreover, for many of its applications, bugs can be critical: money may be lost, for example. These two factors put a significant premium on programs being \emph{correct}, that is, the chain
behaves in the way the author of the code intended, and the participants (and
readers of the code) expect.

\begin{itemize}
\item A type system helps get rid of a large class of bugs (unexpected
  behaviour), and thus promotes correctness. In the future we will explore if a
  type system can be added to \rad{} while still allowing for eval redefinition.
\item Most type systems won't prevent serious problems such as security issues,
  and those that would are likely to be inconvenient for low-risk chains.
  Therefore we will also investigate developing other sorts of tooling around
  \rad{} for verifying correctness. One example we have in mind is a property
  testing tool in which one specifies a schema for valid inputs to a chain, and
  some properties which should be maintained. The tool would then generate many
  simulations in an attempt to find a minimal counter-example to the stated
  properties.
\end{itemize}
